---
title: "STAT5002 Lab7 Worksheet"
subtitle: "Confidence Intervals and Z-test"
author: "© University of Sydney STAT5002"
date: 
output:
  html_document:
    fig_caption: yes
    number_sections: yes
    self_contained: yes
    theme: flatly
    css: 
      # - https://use.fontawesome.com/releases/v5.0.6/css/all.css
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: show
---
<style>
h2 { /* Header 2 */
    font-size: 22px
}
</style>

<style>
h3 { /* Header 3 */
    font-size: 18px
}
</style>

# <i class="fas fa-chart-bar"></i> Same-Sex Marriage Survey

Between 12 September and 7 November 2017 the Australian Government conducted a national postal survey in order to gauge the Australian community's opinion on changing legislation about same-sex marriage.

In the lead-up to the postal survey, various organisations conducted their own polls to infer how the final survey might turn out. [One such poll](https://www.theguardian.com/australia-news/2017/aug/21/most-support-marriage-equality-and-80-plan-to-vote-in-survey-guardian-essential-poll/) returned the following results:

| In favour of proposed change | Against proposed change | Don't know | 
|:----------------------------:|:-----------------------:|:----------:|
|   1036                       | 581                     |   200      |

What can we infer from this about the final results of the official postal survey?

## Q1.3: Confidence interval {-}

- **Determine a Wilson's 99% confidence interval for $p$ based on the observations in the previous part. Hint:** use `binom.confint()`.


```{r}
require(binom)    
# write code here

# sample size
n = 1036 + 581

# sample sum
s = 1036 

# observe sample proportion 
p = s / n

# coMpute CI
CI = binom.confint(s, n, conf.level = 0.99, method = "wilson")

CI
```

- **Perform a "sanity check" that the endpoints of the interval are such that the observed value $\bar x$ is "right on the edge" of a corresponding 99% prediction interval.**

  - You need to identify the appropriate multiplier for a 99% prediction interval
  - Note that the endpoints of the Wilson interval are available via the objects `CI$lower` and
`CI$upper`:

```{r}
# write code here for the confidence interval

# multiplier
qnorm(0.995)

# extracting CI endpoints
CI$lower
CI$upper

# manually computing CI using the lower bound:
CI$lower + c(-1,1) * qnorm(0.995) * sqrt(CI$lower * (1 - CI$lower) / n)

# manually computing CI using the upper bound:
CI$upper + c(-1,1) * qnorm(0.995) * sqrt(CI$upper * (1 - CI$upper) / n)
```

## Q1.4: Hypothesis test {-}

**The observed proportion is greater than 0.5, but is it significantly greater? Answer this question with an appropriate formal hypothesis test.**

* Hypotheses:
$H_0: p = 0.5 $
$H_1: p > 0.5 $

* Assumptions

- poll sample is like a random sample taken with replacement from a box model containing only 0s and 1s, with proportion of 1s equal to some unknown $p$

* Test statistic

```{r}
# write code here for the P-value

z = 2 * sqrt(1617) * (0.64 - 0.5)
z

# p - value

pnorm(z, lower.tail = F)

```
* Conclusion

The p-value less than the significance level of 0.05, meaning there is insufficient evidence of retain $H_0,$ and thus we must reject $H_0.$ Therefore, $p > 0.5$
